---
date: "August 29, 2018"
output: 
    html_document:
        toc: true
        toc_float: true
        toc_depth: 4
        df_print: paged
        code_folding: hide
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# GND Sentinel Site Analyses {.tabset}  


Setup  
-------------------------------------------------------------------------------

### General file setup functions

#### Load libraries and source function file 
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(knitr)
library(broom)
source('000-Functions.R')
```

#### Load data  

a.  csv files: 
    +  *markers_UPDATED* - most recent, correct marker horizon data  
    +  *sets_wide_UPDATED* - most recent, correct pin data    
    +  *NAVD88* - elevations of SET head adapters at each site; generated in 2012? (unsure)  

b.  xlsx: last feldspar application dates (may not need this one?)  

```{r}
markers_raw <- read.csv('../data/markers_UPDATED.csv')
sets_raw <- read.csv('../data/sets_wide_UPDATED.csv')
navd88_setheads <- read.csv('../data/NAVD88.csv')
```

#### Format the data frames - deal with dates; attach site names 

a.  markers:  

```{r}
markers <- markers_raw %>%
    mutate(date = lubridate::mdy(date),
           app_date = lubridate::mdy(ref_date)) %>%
    gather(key = ncore, value = value, A, B, C, D) %>%
    select(SET, MHplot, ncore, date, app_date, value) %>%
    arrange(SET, date, MHplot, ncore) %>%
    attach_site_names()
```
  
  
b.  pins:    

```{r}
pins <- sets_raw %>%
    mutate(date = lubridate::mdy(date)) %>%
    gather(key = npin, value = value, -SET, -date, -arm) %>%
    arrange(SET, date, arm, npin) %>%
    attach_site_names()
```


#### General housekeeping:  

```{r}
sessionInfo()
```




Calculated Rates  {.active}  
------------------------------------------------------------------------------

### Number Crunching and output  

**NOTE: I've rounded everything to 3 digits in the output tables**  

This can be changed if needed. 

Output data frames will be:  

  a.  point measurements of rate (SET: by pin; MH: by MHplot)  
  b.  average rate per SET plot (SET: avg of 36 pin rates; MH: avg of 3 plot rates)  
  c.  average per site (for both, avg of 3 SET plot rates)  <-- this will be output from the ANOVA/Tukey tests though, so may not have to go in a separate table?

### Elevation Change Rates (SETs)  

***

#### By pin

**Linear regression for each pin.** Keeping errors and confidence intervals in here just so we can describe the linear regression model should it ever be necessary. But really we'll only be working with *rate_mm.yr* from here on out.    
```{r, rows.print = 9}
models <- pins %>%
    group_by(SET, arm, npin) %>%
    do(mod = lm(value ~ date, data = .))  

modelcoef <- tidy(models, mod, conf.int=TRUE, conf.level=0.95) # this gives the intercept, slope, and p values, among others; conf.int=TRUE, conf.level=0.95 gives the 95% confidence interval

# get rid of intercept row,
# calculate mm/yr for all the terms,
# round mm/yr to 3 digits,
# get rid of the mm/day stuff
rates_elev_pins <- modelcoef %>%
    filter(term == "date") %>%
    mutate(rate_mm.yr = round(estimate*365.25, 3),
           se_mm.yr = round(std.error*365.25, 3),
           CIlow_mm.yr = round(conf.low*365.25, 3),
           CIhigh_mm.yr = round(conf.high*365.25, 3),
           p.value = round(p.value, 4),
           statistic = round(statistic, 3)) %>%
    select(SET, arm, npin, rate_mm.yr, se_mm.yr, CIlow_mm.yr, CIhigh_mm.yr, statistic, p.value)

modelsummary <- glance(models, mod) # this gives r^2 and adjusted r^2, and p values, among others

rates_elev_pins

```


***
    
#### By SET  

Simple average of the 36 pins per SET.  

```{r}
rates_elev_set <- rates_elev_pins %>%
    group_by(SET) %>%
    summarize(rate = mean(rate_mm.yr, na.rm = TRUE),
              sd_rate = sd(rate_mm.yr, na.rm = TRUE)) %>%
    ungroup() %>%
    attach_site_names()

kable(rates_elev_set, 
      align = 'c',
      digits = 3,
      caption = "Mean and standard deviation of elevation rate change in mm per year by SET")
```

***

#### By site  

Average of 3 SETs per site  

```{r}
rates_elev_site <- rates_elev_set %>%
    group_by(site) %>%
    rename(rates = rate) %>%
    summarize(rate = mean(rates, na.rm = TRUE),
              sd_rate = sd(rates, na.rm = TRUE)) %>%
    ungroup() 

kable(rates_elev_site, 
      align = 'c', 
      digits = 3,
      caption = "Mean and standard deviation of elevation rate change in mm per year by site")
```

***
***

### Accretion (Marker Horizons)  

  


    
    
Statistical Analyses    
-------------------------------------------------------------------------------

### Statistical Analyses  
    
1.  Compare change between sites - ANOVA and Tukey  




Graphics  
-------------------------------------------------------------------------------

### Graphics  

1.	SETs by SET plot and Site (include SLR line)
2.	MH by SET plot and Site  
3.	SETs and MH together by plot (include SLR line)
4.	SETs and MH together by Site. (include SLR line) 
5.  NAVD88 SLR graphic  
    +  make sure the slope for each site is the same as the average of the three SET plots' slopes, or the graph won't match the tables/ANOVA outputs.  
    +  OR construct the line from those statistical outputs, rather than geom_smooth. Probably safer. Only rates in those though, so intercept would be.... average of NAVD88 elevations of three SETS at a site at the beginning of the study?  
